Augmenting Library Development by Mining Usage Data from~Downstream~Dependencies

### 2022-04-29 docs/paper/paper.tex
Restructure paper: extract split up approach, merge implementation
paper: minor phrasing and typesetting improvements
Refactor and fix figure and footnote positioning
Try to fix paper template on CI, second attempt
Supersedes 99750f408845efee3049273a7f16a9deaca543b8.
Fix footnote positioning; use ACM colors
Add evaluation
Revise paper and apply feedback by @scheibel
- Restructure sections
- Shorten things
- Rephrasing
- Improve typesetting
Add paper versions as published in ENASE
Co-authored-by: Willy Scheibel <willy.scheibel@hpi.de>
Co-authored-by: cgcostume <cgdaniel@outlook.com>

\input{template/acm}
\input{template/bibtex}
\input{template/hyphenation}
\input{template/packages}
\input{template/commands}

@@

	\input{title}

	\input{sections/0_abstract}
	\keywords{mining software repositories, ecosystem call graphs, API usage, type analyzer, npm}

	\maketitle

	\input{sections/1_introduction}
	\input{sections/2_related_work}
	\input{sections/3_approach}
	\input{sections/4_data_collection}
	\input{sections/5_usage_mining}
	\input{sections/6_presentation}
	\input{sections/7_evaluation}
	\input{sections/8_conclusion}

	\input{acknowledgements}

	\input{template/bibliography}

	%\appendix

\end{document}
\endinput

>>>>

\documentclass[a4paper,twoside]{article}

\usepackage[T1]{fontenc}
\usepackage{epsfig}
\usepackage{subcaption}
\usepackage{calc}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{multicol}
\usepackage{pslatex}
\usepackage{apalike}
\usepackage[bottom,hang,flushmargin]{footmisc} % added hang,flushmargin
\usepackage{csquotes}

% OWN PACKAGES
\usepackage{microtype}
\usepackage{numprint}
\usepackage{tabularx}
\usepackage{balance}

\usepackage[table]{xcolor}
\usepackage{calc}
\usepackage{tikz}
\usetikzlibrary{calc,positioning,spy,shapes}

\tikzset{thick/.style = {line width = 1.0pt}}

\newlength{\pfull}  \setlength{\pfull}  {2mm}
\newlength{\phalf}  \setlength{\phalf}  {0.5\pfull}

\newlength{\lwffp}   \setlength{\lwffp}   {0.5000\dimexpr(\linewidth-1.0\pfull)}
\newlength{\lwfffp}  \setlength{\lwfffp}  {0.3333\dimexpr(\linewidth-2.0\pfull)}
\newlength{\lwffffp} \setlength{\lwffffp} {0.2500\dimexpr(\linewidth-3.0\pfull)}
\newlength{\lwfffffp}\setlength{\lwfffffp}{0.2000\dimexpr(\linewidth-4.0\pfull)}

\newlength{\twffp}   \setlength{\twffp}   {0.5000\dimexpr(\textwidth-1.0\pfull)}
\newlength{\twfffp}  \setlength{\twfffp}  {0.3333\dimexpr(\textwidth-2.0\pfull)}
\newlength{\twffffp} \setlength{\twffffp} {0.2500\dimexpr(\textwidth-3.0\pfull)}
\newlength{\twfffffp}\setlength{\twfffffp}{0.2000\dimexpr(\textwidth-4.0\pfull)}

\tikzstyle{labelStyle} = [ shape = rectangle, font = \scriptsize, text depth = 0, fill = white, inner sep = 4pt ]
\tikzstyle{labelN}  = [ labelStyle, anchor = north ]
\tikzstyle{labelNW} = [ labelStyle, anchor = north west ]
\tikzstyle{labelNE} = [ labelStyle, anchor = north east ]
\tikzstyle{labelS}  = [ labelStyle, anchor = south ]
\tikzstyle{labelSW} = [ labelStyle, anchor = south west ]
\tikzstyle{labelSE} = [ labelStyle, anchor = south east ]

\tikzstyle{figRC} = [shape = rectangle, rounded corners = 2mm, clip] % figure with Rounded Corners
\tikzstyle{spyRC} = [shape = rectangle, rounded corners = 2mm, %
    white, connect spies, every spy on node/.append style = {thick}] % spy with Rounded Corners
\tikzstyle{figNS} = [inner sep = 0pt, outer sep = 0pt] % figure with No Separators

\newcommand*\rot{\rotatebox{90}}
\newcommand{\rott}[2]{\multicolumn{1}{#1}{\rot{#2}}}

\usepackage[hidelinks]{hyperref}

% Define starred commands
\usepackage{xparse}

% Enumerations
\usepackage[inline]{enumitem}

% Draw figures
% \usepackage{tikz}
% \includegraphs for .tikz files
\usepackage{tikzscale}

% Lay out multiple figures
\usepackage{subcaption}

% Extended control of figure positioning
\usepackage{dblfloatfix}

\usepackage{float}
\newfloat{lstfloat}{htbp}{lop}
\floatname{lstfloat}{Listing}
\def\lstfloatautorefname{Listing} % needed for hyperref/auroref

% Algorithms
\usepackage[noend,noline,boxed]{algorithm2e}
\SetAlCapSkip{1em}
% Python style
\DontPrintSemicolon
\SetStartEndCondition{ }{}{}%
\SetKwIF{If}{ElseIf}{Else}{if}{:}{else if}{else}{end if}%
\SetKwFor{For}{for}{:}{next}%
\SetKw{kwYield}{yield}

% Source code listings
\usepackage{listings}
\usepackage{lstset-typescript}

% Underlined text with linebreaks
\usepackage[normalem]{ulem}

% SI units
\usepackage{siunitx}
\sisetup{input-ignore={,},input-decimal-markers={.}}  % allow for manual thousands separators in numbers
\sisetup{group-separator = {,}}
\DeclareSIUnit\bit{bit}
\DeclareSIUnit\byte{B}

\usepackage{xcolor}

% Convenient references (needs to be loaded after algorithms)
\usepackage[nameinlink]{cleveref}

% Tables
\usepackage{booktabs}
\usepackage{multirow}

\usepackage{makecell}
\renewcommand\theadalign{bc}

\usepackage{threeparttable}

% Bar charts
\usepackage{bchart}

\definecolor{plots-1}{RGB}{228,26,28}
\definecolor{plots-2}{RGB}{55,126,184}

\usepackage{tikz}
\usetikzlibrary{calc,fit,backgrounds,positioning,shapes,snakes,arrows.meta,spy,shapes.multipart,patterns}

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{pgfplotstablefilter}
\pgfplotsset{compat=newest}% es muss mindestens Version 1.8 installiert sein
\usepgfplotslibrary{statistics}

% END OWN PACKAGES

\usepackage{SCITEPRESS}     % Please add other packages that you may need BEFORE the SCITEPRESS.sty package.

% Semantic markup
\newcommand\caps[1]{%
	\textsc{#1}%
}
\NewDocumentCommand\mycode{sm}{%
    \IfBooleanTF#1%
        {{\small\texttt{#2}}}% If a star is seen
        {{\small\texttt{\detokenize{#2}}}}%     If no star is seen
}

% Semantic mathmode markup
\newcommand{\mli}[1]{\mathit{#1}}
\newcommand{\mtx}[1]{\text{#1}}

% Todo comments
\newcommand\TODO[1]{%
    {\leavevmode\bfseries\color{red}#1}%
}

% bchart legends
% CREDITS: https://tex.stackexchange.com/a/379816/221054
\usetikzlibrary{fit}

\makeatletter
\newdimen\legendxshift
\newdimen\legendyshift
\newcount\legendlines
% distance of frame to legend lines
\newcommand{\bclldist}{1mm}
\newcommand{\bclegend}[3][10mm]{%
    % initialize
    \legendxshift=0pt\relax
    \legendyshift=0pt\relax
    \xdef\legendnodes{}%
    % get width of longest text and number of lines
    \foreach \lcolor/\ltext [count=\ll from 1] in {#3}%
        {\global\legendlines\ll\pgftext{\setbox0\hbox{\bcfontstyle\ltext}\ifdim\wd0>\legendxshift\global\legendxshift\wd0\fi}}%
    % calculate xshift for legend; \bcwidth: from bchart package; \bclldist: from node frame, inner sep=\bclldist (see below)
    % \@tempdima: half width of bar; 0.72em: inner sep from text nodes with some manual adjustment
    \@tempdima#1\@tempdima0.5\@tempdima
    \pgftext{\bcfontstyle\global\legendxshift\dimexpr\bcwidth-\legendxshift-\bclldist-\@tempdima-0.72em}
    % calculate yshift; 5mm: heigt of bar
    \legendyshift\dimexpr5mm+#2\relax
    \legendyshift\legendlines\legendyshift
    % \bcpos-2.5mm: from bchart package; \bclldist: from node frame, inner sep=\bclldist (see below)
    \global\legendyshift\dimexpr\bcpos-2.5mm+\bclldist+\legendyshift
    % draw the legend
    \begin{scope}[shift={(\legendxshift,\legendyshift)}]
    \coordinate (lp) at (0,0);
    \foreach \lcolor/\ltext [count=\ll from 1] in {#3}%
    {
        \node[anchor=north, minimum width=#1, minimum height=5mm,fill=\lcolor] (lb\ll) at (lp) {};
        \node[anchor=west] (l\ll) at (lb\ll.east) {\bcfontstyle\ltext};
        \coordinate (lp) at ($(lp)-(0,5mm+#2)$);
        \xdef\legendnodes{\legendnodes (lb\ll)(l\ll)}
    }
    % draw the frame
    \node[draw, inner sep=\bclldist,fit=\legendnodes] (frame) {};
    \end{scope}
}
\makeatother

% Properly spaces tnotes for threeparttable
% CREDITS: https://tex.stackexchange.com/a/108640/221054
% This is like `\tnote`, but uses the space of the argument
\newcommand\stnote[1]{\textsuperscript{\TPTtagStyle{#1}}}

% Alpha numbers for threeparttable tnotes
\makeatletter
\newcommand{\alphtnotetext}[1]{\itshape \@alph{#1}}
\makeatother
\newcommand{\alphtnote}[1]{\stnote{\alphtnotetext{#1}}}

\definecolor[named]{ACMBlue}{cmyk}{1,0.1,0,0.1}
\definecolor[named]{ACMYellow}{cmyk}{0,0.16,1,0}
\definecolor[named]{ACMOrange}{cmyk}{0,0.42,1,0.01}
\definecolor[named]{ACMRed}{cmyk}{0,0.90,0.86,0}
\definecolor[named]{ACMLightBlue}{cmyk}{0.49,0.01,0,0}
\definecolor[named]{ACMGreen}{cmyk}{0.20,0,1,0.19}
\definecolor[named]{ACMPurple}{cmyk}{0.55,1,0,0.15}
\definecolor[named]{ACMDarkBlue}{cmyk}{1,0.58,0,0.21}

\colorlet{paperbackground}{white}
\colorlet{accent1}{ACMBlue}
\colorlet{accent2}{ACMYellow}
\colorlet{accent3}{ACMOrange}
\colorlet{accent4}{ACMRed}
\colorlet{accent5}{ACMLightBlue}
\colorlet{accent6}{ACMGreen}
\colorlet{accent7}{ACMPurple}
\colorlet{accent8}{ACMDarkBlue}

\colorlet{bchart.accent1}{accent1!70!paperbackground}
\colorlet{bchart.accent2}{accent2!70!paperbackground}
\colorlet{bchart.accent3}{accent3!70!paperbackground}
\colorlet{bchart.accent4}{accent4!70!paperbackground}
\colorlet{bchart.accent5}{accent5!70!paperbackground}
\colorlet{bchart.accent6}{accent6!70!paperbackground}
\colorlet{bchart.accent7}{accent7!70!paperbackground}
\colorlet{bchart.accent7}{accent7!70!paperbackground}
\colorlet{bchart.accent8}{accent8!70!paperbackground}

\lstset{frame=none,
	showstringspaces=false,
	columns=flexible,
	basicstyle={\small\ttfamily},
	numbers=none,
	keywordstyle=\textbf,
	stringstyle=\color{blue},
	breaklines=true,
	breakatwhitespace=true,
	tabsize=3
}

\colorlet{punct}{black}
\definecolor{delim}{RGB}{20,105,176}
\colorlet{numb}{magenta!60!black}
\lstdefinelanguage{json}{
    frame=lines,
    literate=
     *{0}{{{\color{numb}0}}}{1}
      {1}{{{\color{numb}1}}}{1}
      {2}{{{\color{numb}2}}}{1}
      {3}{{{\color{numb}3}}}{1}
      {4}{{{\color{numb}4}}}{1}
      {5}{{{\color{numb}5}}}{1}
      {6}{{{\color{numb}6}}}{1}
      {7}{{{\color{numb}7}}}{1}
      {8}{{{\color{numb}8}}}{1}
      {9}{{{\color{numb}9}}}{1}
      {:}{{{\color{punct}{:}}}}{1}
      {,}{{{\color{punct}{,}}}}{1}
      {\{}{{{\color{delim}{\{}}}}{1}
      {\}}{{{\color{delim}{\}}}}}{1}
      {[}{{{\color{delim}{[}}}}{1}
      {]}{{{\color{delim}{]}}}}{1},
}

\lstdefinelanguage{sourcegraph}{
	alsodigit={-},
	morekeywords={
		count,
		file,
		-file,
		select
	},
	literate=*{\\.}{{{\color{accent4}{\textbackslash.}}}}{1},
	moredelim=[is][\textbf]{\ }{:},
	% placeholders
	moredelim=[s][\color{accent1}\itshape]{<}{>},  % TODO: italic does not work
	morestring=[b]
}

% Left-align footnotes in minipages - this reverts a part of the acm style
% CREDITS: https://github.com/borisveytsman/acmart/issues/449#issuecomment-932653376
\makeatletter
\long\def\@mpfootnotetext#1{%
\global\setbox\@mpfootins\vbox{%
  \unvbox\@mpfootins
  \reset@font\footnotesize
  \hsize\columnwidth
  \@parboxrestore
  \protected@edef\@currentlabel
	   {\csname p@mpfootnote\endcsname\@thefnmark}%
  \color@begingroup%\centering - commented out centering footnotes
	\@makefntext{%
	  \rule\z@\footnotesep\ignorespaces#1\@finalstrut\strutbox}%
  \color@endgroup}}
\makeatother

\newcommand\footnoteseparator{\textsuperscript{,}}

\crefname{requirement}{requirement}{requirements}
\crefname{question}{question}{questions}

@@

\title{Augmenting Library Development by Mining Usage Data from~Downstream~Dependencies}

\author{\authorname{Christoph Thiede, Willy Scheibel, Daniel Limberger, and J{\"u}rgen D{\"o}llner}
\affiliation{Hasso Plattner Institute, Digital Engineering Faculty, University of Potsdam}}

\keywords{Mining Software Repositories, Downstream Dependencies, API Usage.}

\inofficialcopyrightnotice{Author's Version. This paper has been published in the Proceedings of the 17th International Conference on Evaluation of Novel Approaches to Software Engineering -- ENASE, 221--232, 2022, and is accessible on \textsc{SciTePress}: \url{https://doi.org/10.5220/0011093700003176}}

\abstract{%
    Software and its dependencies build up a graph where edges connect packages according to their dependencies.
    In this graph, downstream dependencies are all the nodes that depend on a package of interest.
    Although gathering and mining such downstream dependencies allow for informed decision-making for a package developer, there is much room for improvement, such as automation and integration of this approach into their development process.
    This paper makes two contributions:
    \begin{enumerate*}[label=(\roman*)]
        \item We propose an approach for efficiently gathering downstream dependencies of a single package and extracting usage samples from them using a static type analyzer.
		\item We present a tool that allows npm package developers to survey the aggregated usage data directly in their IDE in an interactive and context-sensitive way.
	\end{enumerate*}
	We evaluate the approach and the tool on a selection of open source projects and specific development-related questions with respect to found downstream dependencies, gathering speed, and required storage.
    Our methods return over \numprint{8000} downstream dependencies for popular packages and process about 12 dependencies per minute.
    The usage sample extraction offers high precision for basic use cases.
    The main limitations are the exclusion of unpopular and closed-source downstream dependencies as well as failing analysis when encountering complex build configurations or metaprogramming patterns.
    To summarize, we show that the tool supports package developers in gathering usage data and surveying the use of their packages and APIs within their downstream dependencies, allowing for informed decision-making and improving the resdesign of APIs.
}

\onecolumn \maketitle \normalsize \setcounter{footnote}{0} \vfill

\section{Introduction}
\label{sec:introduction}

The continued proliferation and improvement of open-source software (OSS) platforms such as \emph{GitHub} or \emph{GitLab} and package deployment ecosystems such as \emph{npm}, \emph{pip}, or \emph{NuGet} support and facilitate software development on a global scale.
Open-source development offers many advantages over traditional closed-source development, including voluntary contributions from the open-source community, greater transparency in security-related areas, and a high potential for solution reuse~\cite{saied2018-reusability}.
These solutions are typically organized as packages, each solving an isolated problem and providing a higher-level interface.
Packages can \emph{depend} on existing packages by including them in their manifest, i.e., the \texttt{package.json} file of npm packages or the \texttt{requirements.txt} file of Python projects.
These dependency relations form a large directed graph connecting significant parts of the software world for today's programming language ecosystems.
We define a software ecosystem as a collection of software projects, which are developed and which co-evolve in the same environment.

Despite this connectedness by design, however, the development process of many packages is still characterized by an isolated approach:
While OSS developers commonly submit tickets to or contribute patches against \emph{upstream repositories} that they depend on to solve subproblems, the reverse direction of these edges -- called \emph{downstream dependencies} and in the following referred to as \emph{dependencies} -- is often neglected by package developers.
This deficit can cause a wide range of alignment issues, including poorly suited interfaces~\cite{piccioni2013empirical}, unidentified defects~\cite{wong2017more}, and compatibility problems~\cite{bogart2015breaks}.
Eventually, all these issues impair the capabilities of the global OSS community to build and support high-quality packages.

To tackle these concerns, we propose an approach to extract API usage samples from downstream-dependency repositories of individual packages to support package developers in exploring usages of their packages.
We make these data directly available to package developers by integrating them into an integrated development environment (IDE).

Our approach consists of three steps:
\begin{enumerate*}[label=(\roman*)]
	\item collect downstream dependency repositories from public source code repositories,
	\item mine package usage samples from these dependencies using a type analyzer,
	and \item aggregate and present these usage data to the package developer in its development environment.
\end{enumerate*}
The approach is implemented as a Visual Studio Code extension\footnote{All presented artifacts are available on GitHub: \href{https://github.com/LinqLover/downstream-repository-mining}{https://github.com/LinqLover/downstream-repository-mining}~\cite{thiede2022dowdep}} and evaluated on a selection of open source projects and specific development-centric questions.

The remainder of this paper is structured as follows:
In \cref{sec:related_work}, we summarize existing approaches and in \cref{sec:approach}, we outline the overall conditions for our solution and detail the collection of downstream dependency repositories and subsequent mining of usage data.
In \cref{sec:implementation}, we present an implementation of our approach and outline design decisions for displaying usage samples to the user.
Finally, in \cref{sec:evaluation}, we examine the fitness of our data mining approach and the usability of the display solution.
\Cref{sec:conclusion} concludes this paper.

\section{Related Work}
\label{sec:related_work}
%
Proper tooling design for improving package developers' knowledge about interface usages in dependencies has not yet attracted significant attention in the scientific community.
Nevertheless, programmatic analysis of dependency graphs and API usage mining are already standard techniques in the field of \emph{mining software repositories} (MSR).
Chaturvedi et al. provide a broad overview of existing achievements and ongoing research topics in this field~\cite{chaturvedi2013tools}.
Besides source code repositories, they describe data sources worthwhile to examine, including telemetry data from IDEs, issue trackers, and discussion platforms, and propose different directions for evaluating the retrieved data, e.g., classifying or ranking repositories, analyzing the evolution of projects, studying development communities, but also inspecting the relationships and dependencies between projects.
We further consider related work in the areas of dependency graphs, usage sample extraction, and tooling support.

\paragraph{Searching dependency graphs.}
\label{sec:related_work/dependencies}
%
A common purpose for \emph{analyzing dependency graphs} is to discover transitive upstream dependencies of a project and assess their impact on the stability and vulnerability of the project~\cite{kikas2017structure}.
Similarly, the spreading of security vulnerabilities along the chain of downstream dependencies can be measured~\cite{decan2018impact}.

However, existing research on dependency graphs takes a broad statistical perspective and relies on a large corpus of downloaded software repositories~\cite{abdalkareem2017developers,kikas2017structure,katz2020libraries}.
As opposed to this approach, these capacities are not suitable for scenarios that developers of a single package occasionally perform.
In this case, developers will often use a public \emph{code search} service instead, such as \emph{GitHub} or \emph{Sourcegraph}.
Liu et al. provide a survey of methods and trends in code search tools that include newer approaches such as structural or semantic search queries, code similarity metrics, or machine learning methods~\cite{liu2020opportunities}.

\paragraph{Extracting usage samples.}
\label{sec:related_work/usage_samples}
%
Dependency graphs do not provide information about package usage at sufficient granularity. For example, insights about referencing package classes or methods are missing.
The process of extracting fine-grained information about all references to individual elements of an interface is referred to as \emph{API usage analysis}~\cite{lammel2011large}.
A simple approach is to perform a string search for package names or members in the dependencies~\cite{mileva2010mining}, or to operate on the \emph{abstract syntax tree} (AST) of each parsed dependency to avoid false positive caused by ambivalent identifier names~\cite{qiu2016understanding}.
In addition, one can collect usage data to analyze the historical importance of certain features supported by an API~\cite{sawant2017fine}.

However, the precision of ASTs is limited for \emph{dynamically typed} languages such as JavaScript or Python.
As opposed to static typing, the value type of an expression in a dynamically typed language may be unknown at the compile time of the program.
For this reason, a \emph{type analysis} can help to predict runtime types of all expressions in the AST~\cite{jensen2009type}.
This is usually accomplished by analyzing the control flow of a program and inferring possible types of every variable or function.
Another source for enhancing type information is \emph{call graphs} that are either gathered \emph{statically}, i.e., as a result of a theoretical analysis of the source code, or \emph{dynamically}, i.e., sampled during actual program execution.
While the latter method is better suited to include more contextual information, the former method has the advantage that it can be applied to a more considerable amount of source code without requiring concrete entry points, parameterization, and run-time data.
Many solutions exist that analyze the structure of programs and extract relevant information to build call graphs:

Collard et al. propose an infrastructure called \caps{srcML} that aims to create a unified representation of multilingual source code snippets for arbitrary analysis purposes, including the construction of call graphs~\cite{collard2013srcml}.
Another solution utilizes an island parser to build polyglot call graphs~\cite{bogar2018lightweight}.
Furthermore, Antal et al. compare several call graph generators for JavaScript and emphasize that static call graph generators have limited precision by design for languages supporting dynamic typing or meta-programming mechanisms~\cite{antal2018static}.

To merge dependency graphs and call graphs, ecosystem call graphs can be constructed by applying call graphs to entire ecosystems that cross repository boundaries.
Many approaches apply this concept to different ecosystems, often aiming to conduct fine-grained impact analysis for security vulnerabilities~\cite{hejderup2018software,boldi2020fine,wang2020empirical,hejderup2021praezi,keshani2021scalable,nielsen2021modular}.
Another approach is the \emph{precise code intelligence} feature of Sourcegraph, which makes code references explorable across repository boundaries for repositories that provide a metadata file in the \emph{Language Server Index Format}.\footnote{\url{https://docs.sourcegraph.com/code_intelligence/explanations/precise_code_intelligence}}

After collecting these usage data, additional processing is possible to extract general usage information from the extensive raw data.
Next to basic grouping and counting operations, such aggregations can be built by detecting popular \emph{usage patterns} of API features in the downstream dependencies.
Zhong et al. propose such a framework, that, for instance, finds sequences of API members that are invoked frequently and even uses these patterns for guiding API users by giving them recommendations~\cite{zhong2009mapo}.
Hanam et al. pursue a different goal in their tool that helps package developers assess the impact of breaking API changes on the functionality of downstream dependencies~\cite{hanam2019aiding}.

\paragraph{Presentation of results.}
\label{sec:related_work/presentation}

To establish an adequate developer experience, any collected usage data still need to be presented suitably.
Due to the hierarchical structure of source code, a common approach for this is a hierarchical and navigatable representation of the collected call tree.
An early form of this representation has been invented as the interactive \enquote{message set} tool for browsing senders and implementors in an object-oriented system for the Smalltalk programming environment \cite{goldberg1984smalltalk}.
Modern alternatives include the \emph{Stacksplorer} \cite{karrer2011stacksplorer} or \emph{Blaze} \cite{kramer2012blaze} that list next to the currently focused method other methods which are adjacent to this method in the call graph.
With a focus on exploring references to API members, De Roover et al. propose a set of additional views, including hierarchical lists and word clouds for highly referenced identifiers \cite{de2013multi}.
Focusing on exploring APIs in a non-code-centric way, Hora and Valente propose a dashboard that allows developers to browse existing APIs and track possibly breaking changes in their interfaces \cite{hora2015apiwave}.

\section{Downstream Dependency Mining}
\label{sec:approach}
%
Our approach to improve developers' knowledge about the usage of their packages is based on the automatic gathering of downstream dependency information and integration into their development environment. This approach assumes a target package and focuses on showing a developer actual usage of parts of their package and, in particular, answering the following questions:
In doing so, we identified three questions from package developers about the usage of their packages:
%
\begin{enumerate}[label=Q\arabic*]
	\item \label[question]{q1} Which dependencies are using the target package, and what problems do they try to solve with them?
	\item \label[question]{q2} By how many dependencies is a particular package member being used?
	\item \label[question]{q3} In which contexts and constellations is a particular package member being used?
\end{enumerate}
%
For an effective integration into the development process, we pose three non-functional requirements:
%
\begin{enumerate}[label=R\arabic*]
	\item \label[requirement]{req1} The tool is ready to use out of the box and requires little configuration (less than 5 minutes for a new user).
	\item \label[requirement]{req2} The tool consumes sufficiently few computational resources to run on a single machine but still delivers interactive response times according to Shneiderman's requirement to frequent tasks~\cite{shneiderman2010designing}.
	\item \label[requirement]{req3} The tool blends in with the usual workflow of package developers at the best possible rate.
\end{enumerate}
%
On a conceptual level, we separate the approach into the collection of downstream dependency projects and mining of usage samples based on a target package.

\begin{algorithm*}[t]
	\caption{Extraction of usage samples.}\label{alg:usage_mining}

	\KwIn{\\\Indp
		$\mli{package}$: target package \\
		$\mli{dependencies}$: downstream dependencies}
	\KwOut{usage samples (set of strings)}
	\;
	\For{$\mli{dependency} \in \mli{dependencies}$}{
		$\mli{asf} \gets \mtx{parse}(\mli{dependency} \cup \mli{package})$\;
	  	$\mtx{annotate\_types}(\mli{asf})$\;
		\For{$\mli{ast} \in \mli{asf}$}{
			\For{$\mli{node} \in \mli{dfs}(\mli{ast})$}{
				\For{$\mli{pattern} \in \mli{patterns}$}{
					\If{$\mli{pattern}.\mtx{matches}(\mli{node}) \wedge\, \mli{package}.\mtx{declares}(\mli{pattern}.\mtx{getType}(\mli{node}))$}{
						\kwYield{$\mli{node}.\mtx{text}$}
					}
				}
			}
		}
	}
\end{algorithm*}

\subsection{Downstream Dependency Collection}
\label{sec:approach/dependency_collection}

In the first step, a list of repositories that depend on the target package has to be assembled.
As mentioned in \cref{sec:related_work/usage_samples}, many approaches start with a large downloaded corpus of unspecific repositories from the ecosystem, which they then iterate over to filter the relevant repositories.
However, a major drawback of this approach is the high resource demand for both creating and traversing this corpus.
This approach is suited for analyzing repositories on a large scale but is not conform to \cref{req1} because of its large footprint in terms of computational power, memory, and time.
As an alternative, we have decided to apply filtering to the set of repositories before downloading a subset of it to the developers' machine.
For that, we use two types of data sources that are available as public web platforms:

\begin{enumerate}[label=(\roman*)]
	\item Package repositories that maintain a doubly-connected edge list of interdependent packages in an ecosystem.
	\item Code search engines that index the source code of many repositories from OSS platforms. Using these search engines, we can query all repositories that declare a dependency on the target package in their package manifest file.
\end{enumerate}

\subsection{Usage Sample Mining}
\label{sec:approach/usage_mining}

Having downloaded the selected downstream dependency repositories, the next step is to extract usage samples for the target package from each dependency repository.
Our goal is to gather fine-grained source code excerpts that reference individual identifiers exposed by the target package.
The complete procedure is displayed in \cref{alg:usage_mining}.

%
To do so, the source code of every dependency repository as well as the source code of the target package are parsed, each into a separate \emph{abstract syntax forest} (ASF, i.e., a set of ASTs).
In a second step, static type analysis against each dependency ASF is performed together with the target package's ASF.
The results of the type analysis are attached to the ASF so that every identifier is annotated with a type symbol that links to the declaration of the type for the identifier.
For instance, after this step, every variable node contains a link to the assignment node of this variable, and every function call expression contains a link to the definition of this function.
The operating principle of this type analysis depends on the kind of programming language.
In statically typed languages, type symbols can be retrieved from declarations, whereas in dynamically typed languages, a control flow analysis will be required to identify the origin of every identifier's type.

In the final step, all nodes whose type is declared in the target package are collected from each ASF.
To identify these links, we use a set of language-specific patterns for AST subtrees that constitute a usage expression.
In particular, each of these patterns expects a node containing a link to an identifier declaration (see \cref{fig:approach/usage_mining/patterns}).

\begin{figure*}
	\begin{minipage}{\linewidth}
		\colorlet{lowlight}{black!40}
		\newcommand\textlowlight[1]{\textcolor{lowlight}{#1}}
		\newcommand\texthighlight[1]{\textbf{\textcolor{accent1}{#1}}}
		\begin{center}
			\begin{subfigure}[t]{.30\linewidth}
				\begin{tikzpicture}
                	\node {CallExpression}
                		child {node [accent1] {\textbf{identifier}}}
                		child [lowlight] {node [lowlight, yshift = -0.5cm] {typeArguments}}
                		child [lowlight] {node [lowlight] {arguments}};
                \end{tikzpicture}
                \caption[LoF entry]{
                	Node pattern for a TypeScript function call, such as in:

                	\mycode*{\textlowlight{result = }\uline{\texthighlight{fun}<\textlowlight{T1}, \textlowlight{T2}>(\textlowlight{arg1}, \textlowlight{arg2})}\textlowlight{;}}
                }

			\end{subfigure}
			\hfill
			\begin{subfigure}[t]{.30\linewidth}
				\begin{tikzpicture}
                	\node {PropertyAccessExpression}
                		child {node [accent1] {\textbf{identifier}}}
                		child [white] {node [yshift = -0.5cm] {\phantom{node}}}  % ensure same height as sibling figures
                		child [lowlight] {node [lowlight] {name}}
                		;
                \end{tikzpicture}
                \caption[LoF entry]{
                	Node pattern for a JavaScript property access, such as in:

                	\mycode*{\textlowlight{return }\uline{\textlowlight{obj}.\texthighlight{prop}}\textlowlight{;}}
                }

			\end{subfigure}
			\hfill
			\begin{subfigure}[t]{.30\linewidth}
				\begin{tikzpicture}
                	\node [accent1] {\textbf{JsxOpeningExpression}}
                		child [lowlight] {node [lowlight] {tagName}}
                		child [lowlight] {node [lowlight, yshift = -0.5cm] {typeArguments}}
                		child [lowlight] {node [lowlight] {attributes}};
                \end{tikzpicture}
                \caption[LoF entry]{
                	Node pattern for a JSX opening element (as supported in React or TypeScript), such as in:

                	\mycode*{\textlowlight{elem = }\uline{<\texthighlight{Button} \textlowlight{color}="\textlowlight{blue}"> \textlowlight{Submit</Button>}}\textlowlight{;}}
                }
			\end{subfigure}

            \vspace{\baselineskip}
			\caption{AST patterns for example JavaScript/TypeScript expressions.
				The \texthighlight{highlighted} node contains the link to the declaration of the referenced identifier.
			}
			\label{fig:approach/usage_mining/patterns}
		\end{center}
	\end{minipage}
\end{figure*}

\section{Prototype}
\label{sec:implementation}

In our prototype, we focus on analyzing JavaScript projects for node.js.
This decision is motivated by the large prevalence of JavaScript\footnote{GitHut 2.0: A small place to discover languages in GitHub. \url{https://madnight.github.io/githut/\#/pushes}} as well as the first-rate extent of the npm ecosystem which has been unsurpassed since 2015\footnote{Modulecounts: \url{http://www.modulecounts.com/}}.
The prototype instantiates the downstream dependency collection and the usage sample mining and allows for the integration of the results into the development environment.

\subsection{Downstream Dependency Collection}
\label{sec:implementation/dependency_collection}

The de-facto standard package repository manager for node.js is npm.
To fetch downstream dependency packages from the npm registry, we use the npm package \texttt{npm-dependants} which scrapes the dependents list from npmjs.com as there does not exist a publicly available API for these data.
To keep the footprint of our implementation small, only a part of the dependents list is scraped.
Then, basic metadata for each found package is requested from the npm registry, including the \emph{tarball} URL of the latest package version, using the package \mycode{package-metadata}.
Finally, each tarball is downloaded and extracted using the package \mycode{download-package-tarball}.

\begin{lstfloat}[t]
\begin{lstlisting}[language=typescript,%json,
	label=sec:implementation/package.json,
	caption={Example \texttt{package.json} file for the \texttt{ripple-lib} package specifying multiple dependencies (truncated). The list of dependencies that are extracted from this example contains \texttt{bignumber.js}, \texttt{https-proxy-agent}, and \texttt{jsonschema}.}]
{
	"name": "ripple-lib",
	"version": "1.10.0",
	"description": "An API for the XRP Ledger",
	"dependencies": {
		"bignumber.js": "^9.0.0",
		"https-proxy-agent": "^5.0.0",
		"jsonschema": "1.2.2"
	}
}
\end{lstlisting}
\end{lstfloat}

For the second method -- based on a code search engine --, we have identified Sourcegraph\footnote{\url{https://sourcegraph.com/}} as a promising platform.
npm packages specify their dependencies in a \mycode{package.json} metadata file (see \cref{sec:implementation/package.json}).
To find all packages depending on a certain package, we can search all available \mycode{package.json} files for a reference to this package identifier.
We refine our search query with some additional constraints to filter out package copies from \mycode{node_modules} folders that downstream developers have pushed accidentally, and to limit the number of results retrieved.
We submit this search query to the publicly available GraphQL API of Sourcegraph\footnote{\url{https://docs.sourcegraph.com/api/graphql}} by using the npm package \mycode{graphql-request}.
After retrieving a search result, we download a snapshot of the full repository by using the package \mycode{download-git-repo}.

Additionally, we enrich the metadata of each package with a few metrics indicating its popularity from the associated GitHub repository if specified.
The GitHub API is accessed via the official \emph{octokit} library\footnote{\url{https://octokit.github.io/}}.

\subsection{Usage Sample Mining}
\label{sec:implementation/usage_mining}

For parsing the dependencies, we use the \emph{TypeScript Compiler API}\footnote{\url{https://github.com/Microsoft/TypeScript/wiki/Using-the-Compiler-API}}.
TypeScript is an extension of JavaScript that adds strong typing; typing is based on a combination of implicit type inference and explicit type declarations.
As a side benefit, our implementation is also able to analyze packages implemented in true TypeScript.

The TypeScript compiler's \emph{binder} component annotates every node of the parsed program with a type symbol if it can resolve one.
After performing a type analysis of the dependency by instrumenting the binder, we collect all nodes from the type-annotated ASF that refer to the target package,
matching each node against a list of defined patterns (similar to \cref{fig:approach/usage_mining/patterns}).
A lookup of the node's type symbol's declaration is performed based on the syntax kinds of the node, its ancestors, and its descendants, and then compared with the source directory of the target package.
